ms_supplementary

### 2018-01-26 man/ms_1_material_methods.Rmd
update

The long-term growth response was analyzed using basal area increment (hereafter BAI, $\mathrm{cm^2 \cdot year^{-1}}$). BAI represents a more accurate indicator of growth than ring-width, since it removes variation in growth attributable to increasing stem circumference [@Biondi2008]. We used measured dbh and raw ring-widths to compute BAI by subtracting twice the annual ring width from the annual diameter, starting from the measured diameter outside the bark [@Piovesan2008]. We used the following equation:

@@

### Disturbance analyses 
Disturbance chronologies were built using tree-ring width to identify abrupt and sustained increases (release event from competition) or decreases (supressions) in radial growth [@Nowacki1997] as indirect estimates of possible disturbance events (e.g. logging) in the past. By using a temporal window of reasonable length this method is intended to filter out the response to short-term changes in temperature and precipitation and minimizes the long-term growth trends [@Nowacki1997; @Fraver2005].

@@

$$\mathrm{\% GC} = \left [ \frac{\left ( M1 - M2 \right )}{M2}\right] \times 100 $$
where $M1$ is preceding 10-year median and $M2$ is subsequent 10-year median. We used medians since they are more robust estimator of central tendency than means [@Rubino2004; @Camarero2011]. Site disturbance chronologies were constructed by averaging the individual disturbances series annually. To separate growth peaks produced by disturbance events and expressing stand-wise disturbances from those by climate, we considered a threshold of 25 % (DUDA) of GC and more than 50% of the individual trees displaying the same growth changes was considered a stand-wise disturbance.

@@

We computed the values of these indices for tree growth (BAI) and greenness during each drought event. We considered 2005 and 2012 as singles drought events. The predrought and postdrought values of each target variable (*i.e.*: BAI or EVI) were computed as the mean value during a period of three years before and after the drougth event respectively. A period of three years was chosen because we found similar results comparing periods of two, three and four years (Figure S3) and this length was used in other studies [Gazol2017].

>>>>

The long-term growth response was analyzed using basal area increment (hereafter BAI, $\mathrm{cm^2 \cdot year^{-1}}$). BAI represents a more accurate indicator of growth than ring-width, since it removes variation in growth attributable to increasing stem circumference after 30-40 years of juvenile increasing growth [@Biondi2008]. We used measured dbh and raw ring-widths to compute BAI by subtracting twice the annual ring width from the annual diameter, starting from the measured diameter outside the bark [@Piovesan2008]. We used the following equation:

@@

### Disturbance analyses 
Disturbance chronologies were built using tree-ring width to identify abrupt and sustained increases (release events from competition) or decreases (supressions) in radial growth [@Nowacki1997] as indirect estimates of possible disturbance events (e.g. logging) in the past. By using a temporal window of reasonable length this method is intended to filter out the response to short-term changes in temperature and precipitation and minimizes the long-term growth trends [@Nowacki1997; @Fraver2005].

@@

$$\mathrm{\% GC} = \left [ \frac{\left ( M1 - M2 \right )}{M2}\right] \times 100 $$
where $M1$ is preceding 10-year median and $M2$ is subsequent 10-year median. We used medians since they are more robust estimator of central tendency than means [@Rubino2004; @Camarero2011]. Site disturbance chronologies were constructed by averaging the individual disturbances series annually. To separate growth peaks produced by disturbance events and expressing stand-wise disturbances from those by climate, we considered a threshold of 50 % (DUDA) of GC and more than 50% of the individual trees displaying the same growth changes was considered a stand-wise disturbance.

@@

We computed the values of these indices for tree growth (BAI) and greenness during each drought event. We considered 2005 and 2012 as singles drought events. The predrought and postdrought values of each target variable (*i.e.*: BAI or EVI) were computed as the mean value during a period of three years before and after the drougth event respectively. A period of three years was chosen because we found similar results comparing periods of two, three and four years (Figure S3) and this length was used in other studies [@Gazol2017].

### 2018-01-26 man/ms_supp.Rmd
update

```{r}
library("tidyverse")
library("here")
library("dplyr")
library("ggrepel")
library("scales")

library("grid")
library("gridExtra")
library("patchwork")

```

@@

## Supplementary 1

Listado de pkgs usados

## Suplementary 2

## Figure S2

@@

Figure S2. Spatial distribution of the EVI-annual trends for the period 2000-2016. Trends ($\tau$) with the most positive values (green) indicate positive trend.

## Figure S3

@@

Figure S3. Percentage of pixels showing browning, greenning or no-changes during the 2005 and 2012 droguht events according to EVI standardized anomalies.

>>>>

```{r}
library("tidyverse")
library("here")
library("dplyr")
library("ggrepel")
library("scales")
library("grid")
library("gridExtra")
library("patchwork")
library("ggcorrplot")
```

@@

## Figures

**Figure S1**. Spatial distribution of the EVI-annual trends ($\tau$) for the period 2000-2016. Negative values (*red* values) indicate negative trends and positive values (*green* values) indicate positive trends.

@@

**Figure S2.** Percentage of pixels showing browning, greenning or no-changes during the 2005 and 2012 droguht events according to EVI standardized anomalies.

@@

**Figure S3.** Correlation

```{r}

corBai <- read.csv(file=here("/out/correla_resilience/", "correla_window_size.csv"), header = TRUE)
corEvi <- read.csv(file=here("/out/correla_resilience/", "correla_window_size_evi.csv"), header = TRUE)
corRwi <- read.csv(file=here("/out/correla_resilience/", "correla_window_size_rwi.csv"), header = TRUE)

generaMatrix <- function(cordf, variable){ 
  m <- cordf %>% filter(var == variable) %>% 
    dplyr::select(-var, -window_size) %>% 
    mutate(w1 = c(2, 2, 3), 
           w2 = c(3,4,4)) 
  mf <- m %>% 
    dplyr::select(r2, w2, w1) %>% rename(w1 = w2, w2 = w1)

  df <- m %>% rbind(mf) %>% 
    add_row(r2 = 1, w1=2, w2=2) %>% 
    add_row(r2 = 1, w1=3, w2=3) %>% 
    add_row(r2 = 1, w1=4, w2=4)

    df <- df %>% 
    mutate(w1 = case_when(
      w1 == 2 ~ 'ws2',
      w1 == 3 ~ 'ws3',
      w1 == 4 ~ 'ws4'),
      w2 = case_when(
      w2 == 2 ~ 'ws2',
      w2 == 3 ~ 'ws3',
      w2 == 4 ~ 'ws4'))

      cor_df <- as.data.frame.matrix(xtabs(r2 ~., df))

return(cor_df)
}

cor_rt <- generaMatrix(corBai, 'rt')
cor_rc <- generaMatrix(corBai, 'rc')
cor_rs <- generaMatrix(corBai, 'rs')

cor_rte <- generaMatrix(corEvi, 'rt')
cor_rce <- generaMatrix(corEvi, 'rc')
cor_rse <- generaMatrix(corEvi, 'rs')

g_rt <- ggcorrplot(cor_rt, type='upper', lab=TRUE, title='BAI Rt', show.legend = FALSE)
g_rc <- ggcorrplot(cor_rc, type='upper', lab=TRUE, title='BAI Rc', show.legend = FALSE)
g_rs <- ggcorrplot(cor_rs, type='upper', lab=TRUE, title='BAI Rs', show.legend = FALSE)
g_rte <- ggcorrplot(cor_rte, type='upper', lab=TRUE, title='EVI Rt', show.legend = FALSE)
g_rce <- ggcorrplot(cor_rce, type='upper', lab=TRUE, title='EVI Rc', show.legend = FALSE)
g_rse <- ggcorrplot(cor_rse, type='upper', lab=TRUE, title='EVI Rs', show.legend = TRUE)

(g_rte + g_rce + g_rse) / (g_rt + g_rc + g_rs) + plot_layout(nrow=2) + plot_annotation(tag_levels = c('a'))
```
